INGRID: Creating Languages in MPS from ANTLR Grammars

### 2016-10-13 chapters/language_definition.tex
Fixes & comments for section 3

During our work on this project, we also discovered/encountered challenges/issues related to input grammars --- intermediate rules in the syntax hierarchy that allow humans to understand the grammar better (structure aspect/component), whitespace (editor aspect).
There is a challenge/issue that we have discovered during our work on this project (in the process of our own implementation), and that is related to construction of the structure aspect.
Grammars are written by humans and therefore contain some structures (intermediary layers) inside of them that help a human brain to understand it better.
However, these structures add unnecessary levels to its syntax hierarchy which lead to problems when using this language inside MPS.
We describe/discuss these challenges in more detail (also using examples) and our solutions to the challenges inside the respective following subsections.

@@

\todo{Pouzit jazyk SimpleXML na running exaple.}

@@

\begin{figure*}[ht]
\centering
\begin{framed}
\begin{alltt}
	\textbf{grammar} \textit{SimpleXML};
	\antlrparserrule{document}  : \antlrparserrule{prolog}? \antlrparserrule{comment}? \antlrparserrule{element} ;
	\antlrparserrule{prolog}    : \antlrliteral{<?xml } \antlrparserrule{attrib}* \antlrliteral{?>} ;
	\antlrparserrule{comment}   : \antlrliteral{<!--} \antlrlexerrule{TEXT} \antlrliteral{-->} ;
	\antlrparserrule{element}   : \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{>} \antlrparserrule{content}* \antlrliteral{</} \antlrlexerrule{Name} \antlrliteral{>} | \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{/>} ;
	\antlrparserrule{attrib}    : \antlrlexerrule{Name} \antlrliteral{="} \antlrlexerrule{TEXT} \antlrliteral{"} ;
	\antlrparserrule{content}   : \antlrlexerrule{TEXT} | \antlrparserrule{element} | \antlrparserrule{comment} | \antlrlexerrule{CDATA} ;
	\antlrlexerrule{Name}      : \antlrlexerrule{NameStartChar} \antlrlexerrule{NameChar}* ;
	\textbf{fragment}
	\antlrlexerrule{DIGIT}     : \antlrregex{[0-9]} ;
	\textbf{fragment}
	\antlrlexerrule{NameChar}  : \antlrlexerrule{NameStartChar} | \antlrliteral{-} | \antlrliteral{_} | \antlrliteral{.} | \antlrlexerrule{DIGIT} ;
	\textbf{fragment}
	\antlrlexerrule{NameStartChar} : \antlrregex{[:a-zA-Z]} ;
	\antlrlexerrule{TEXT}      : \antlrregex{~[<"]*} ;
	\antlrlexerrule{CDATA}     : \antlrliteral{<![CDATA[} \antlrregex{.*?} \antlrliteral{]]>} ;
\end{alltt}
\end{framed}
\caption{SimpleXML grammar}
\end{figure*}

@@

\begin{center}
	\texttt{([:a-zA-Z])(([:a-zA-Z])|(-)|({\_})|(.)|([0-9]))*}
\end{center}

>>>>

During our work on this project, we also discovered/encountered challenges/issues related to input grammars --- intermediate rules in the syntax hierarchy that allow humans to understand the grammar better (structure aspect/component), whitespace (editor aspect).
There is a challenge/issue that we have discovered during our work on this project (in the process of our own implementation), and that is related to construction of the structure aspect.

Grammars are written by humans and therefore contain some structures (intermediary layers) inside of them that help a human brain to understand it better.
However, these structures add unnecessary levels to its syntax hierarchy which lead to problems when using this language inside MPS.
\PV{Dalsi veta asi zbytecna}
We describe/discuss these challenges in more detail (also using examples) and our solutions to the challenges inside the respective following subsections.

@@

\todo{Pouzit jazyk SimpleXML na running exaple.}
\PV{Dalsi veta asi prijde niz?}

@@

\begin{figure*}[ht]
\centering
\begin{framed}
\begin{alltt}
	\textbf{grammar} \textit{SimpleXML};
	\antlrparserrule{document}  : \antlrparserrule{prolog}? \antlrparserrule{comment}? \antlrparserrule{element} ;
	\antlrparserrule{prolog}    : \antlrliteral{<?xml } \antlrparserrule{attrib}* \antlrliteral{?>} ;
	\antlrparserrule{comment}   : \antlrliteral{<!--} \antlrlexerrule{TEXT} \antlrliteral{-->} ;
	\antlrparserrule{element}   : \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{>} \antlrparserrule{content}* \antlrliteral{</} \antlrlexerrule{Name} \antlrliteral{>} | \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{/>} ;
	\antlrparserrule{attrib}    : \antlrlexerrule{Name} \antlrliteral{="} \antlrlexerrule{TEXT} \antlrliteral{"} ;
	\antlrparserrule{content}   : \antlrlexerrule{TEXT} | \antlrparserrule{element} | \antlrparserrule{comment} | \antlrlexerrule{CDATA} ;
	\antlrlexerrule{Name}      : \antlrlexerrule{NameStartChar} \antlrlexerrule{NameChar}* ;
	\textbf{fragment}
	\antlrlexerrule{DIGIT}     : \antlrregex{[0-9]} ;
	\textbf{fragment}
	\antlrlexerrule{NameChar}  : \antlrlexerrule{NameStartChar} | \antlrliteral{-} | \antlrliteral{\_} | \antlrliteral{.} | \antlrlexerrule{DIGIT} ;
	\textbf{fragment}
	\antlrlexerrule{NameStartChar} : \antlrregex{[:a-zA-Z]} ;
	\antlrlexerrule{TEXT}      : \antlrregex{~[<"]*} ;
	\antlrlexerrule{CDATA}     : \antlrliteral{<![CDATA[} \antlrregex{.*?} \antlrliteral{]]>} ;
\end{alltt}
\end{framed}
\caption{SimpleXML grammar}
\end{figure*}

@@

\begin{center}
	\texttt{[:a-zA-Z]([:a-zA-Z]|\textbackslash-|{\_}|\textbackslash.|[0-9])*}
\end{center}
